---
title: "Exploring Player Survival Rates"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse, jsonlite, lubridate, googlesheets4)
cbGames <- read_sheet("19ONeLaNEHbVALB2iOUG93FPjP7wP06SXQC3V9JpSZNc",sheet=1)
cbPlayers <- read_sheet("19ZMSWKDTOG_KL4YSFfO0_Uu4eZnZ7E0xRjM2kgYTFOg",sheet=1)
```


With the data currently available, it is difficult to directly infer or correlate player performance with game outcomes. There is a near total absence of individual performance statistics, and there are a huge number of slightly awkward variables which change from session to session which make it difficult to attribute successes or failures to individuals or individual changes. Even when changes are evident (different class or ship, e.g) the sample size for that change can be ridiculously small, or it can be accompanied by so many other team-wide changes that it could easily be noise rather than cause.

Alongside winrate, which is team-wide so less easily attributed individually, I have survival as a genuine individual performance metric. So, the question arises: can I draw some kind of conclusion about player performance or game impact based on the game outcome and the player's survival? Are there some players who consistently die, consistently die when we win, or consistently survive when we lose? If not individual players, can anything be inferred from whether specific classes or specific ships consistently survive or die in wins/losses?

Obviously, there should be a correlation between players surviving and winning, since everyone dying is a loss condition, but I don't know how strong that correlation will be, or whether it does/does not exist with regards to specific individuals as well as team aggregates.

```{r grab current season, include=FALSE}
games <- cbGames %>% filter(Season==15)
players <- cbPlayers %>% filter(GID %in% games$GID, Clan=="RUMR")
```

The first thing to look at is, quite simply, survival rates versus winrates:

```{r}

playerRates <- players %>% group_by(Name) %>% summarise(SurRate = mean(as.numeric(Survived)),WinRate=mean(Result))

playerRates %>%
  ggplot(aes(x=SurRate,y=WinRate,label=Name))+
  scale_color_manual(values=c("TRUE"="cornflowerblue","FALSE"="darkred"))+
  geom_point(aes(colour=SurRate>0.5))+
  geom_text(hjust=1.2,size=2)+
  geom_line(linetype="dotdash",aes(y=0.5, colour="lightgrey",alpha=0.5))+
  geom_line(linetype="dotdash",aes(x=0.5, colour="lightgrey",alpha=0.5))+
  theme_classic()+
  guides(color=guide_legend(title=""))+
  ggtitle("WR vs Survival S15")+
  ylab("Winrate")+
  xlab("Survival Rate")+
  geom_smooth(method='lm', formula= y~x)+
  theme(legend.position = "none")
```

This isn't bad, actually: there is a non-zero visible correlation between winrate and individual survival rate, but it is not especially strong. This will also somewhat fail to catch weirdo outliers - those with especially few or many games.

I do need to survive better, though, but I get sloppy when games are won.. and probably when they aren't.

So. Followons: Survival faceted by win/loss and by ship class.

```{r}

playerRates <- players %>% group_by(Name,Result) %>% summarise(SurRate = mean(as.numeric(Survived)),.groups = 'drop')

playerRates %>%
  ggplot(aes(x=Name,y=SurRate,fill=SurRate>0.5,label=Name))+
  scale_fill_manual(values=c("TRUE"="cornflowerblue","FALSE"="darkred"))+
  geom_col()+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  guides(color=guide_legend(title=""))+
  ggtitle("Survival faceted by Winrate")+
  ylab("Survival Rate")+
  xlab("")+
  theme(legend.position = "none")+
  facet_grid(vars(Result))
```

So, there is obviously going to be a correlation between winning and surviving - even absolute sausages like me survive far more when they win. Some weirdos: Awetec, Bemse, Nephstar, OIFson, Vivi_Ornitier all have >50% survival when losing, with some VERY high - Bemse is probably within a 'normal' range of surviving just slightly over 50% instead of slightly under. With the exception of OIFson, everyone with a poor survival rate while winning has a *better* survival rate while winning than while losing.

So: a few of these are probably very low sample sizes. Easy to check:

```{r}
players %>% filter(Name %in% c("OlFson","Bemse","Awetec","Nephstar","Vivi_Ornitier")) %>% group_by(Name,Result) %>% summarise(SurRate = mean(as.numeric(Survived)), N=n(),.groups = 'drop')
```

So: Awetec and OlFson are just sample size issues. Bemse survives A LOT, which could mean he is very good, or could be a bit overpassive. Ditto Nephstar. Vivi is low sample, again, and 100% survival rates. Maybe there is something to do with the ships people are playing, maybe there is nothing to see here.

Bemse, Nephstar, and Vivi - ship classes:
```{r}
players %>% filter(Name %in% c("Bemse","Nephstar","Vivi_Ornitier")) %>% group_by(Name,Result,Class) %>% summarise(SurRate = mean(as.numeric(Survived)), N=n(),.groups = 'drop')
```

I would have expected both BBs and CAs to die quite often when we lose games: losing with ships alive - especially beefy ones - implies passivity.
```{r}
players %>% filter(Name %in% c("Bemse","Nephstar","Vivi_Ornitier")) %>% group_by(Name,Result,Ship) %>% summarise(SurRate = mean(as.numeric(Survived)), N=n(),.groups = 'drop')
```

No firm conclusions, but it is interesting at least.

So, backing out to everyone again: is there a survival difference when splitting by ship classes?

```{r}
playerRates <- players %>% group_by(Name,Class) %>% summarise(SurRate = mean(as.numeric(Survived)),WinRate=mean(Result),N=n(),.groups="drop") %>% filter(N>2)

playerRates %>%
  ggplot(aes(x=SurRate,y=WinRate,label=Name))+
  scale_color_manual(values=c("TRUE"="cornflowerblue","FALSE"="darkred"))+
  geom_point()+
  geom_text(hjust=1.2,size=2)+
  geom_line(linetype="dotdash",aes(y=0.5, colour="lightgrey",alpha=0.5))+
  geom_line(linetype="dotdash",aes(x=0.5, colour="lightgrey",alpha=0.5))+
  theme_classic()+
  guides(color=guide_legend(title=""))+
  ggtitle("Winrate and Survival by Class")+
  ylab("Winrate")+
  xlab("Survival Rate")+
  theme(legend.position = "none")+
  geom_smooth(method='lm', formula= y~x)+
  facet_wrap(vars(Class))
```

If I filter out people with fewer than 2 games (a low bar, but here we are) then we see a relatively weak but consistent positive correlation between survival rate and winrate. The correlation is weakest with DDs, which both makes perfect sense and is relatively counterintuitive: the highest risk-reward class is definitely more valuable as the game goes on, but if they aren't doing anything, they aren't creating any value, so playing 'too safe' is as bad as being 'too aggressive' - it is about generating value not surviving per se.

So: I need to consider my survival more, regardless, and if you're going to die, die for a reason.

Since the current season dataset is quite small, I'll have a look at the last two seasons individually, with the same visualisations, filtering out players with five or fewer games:

```{r}
games14 <- cbGames %>% filter(Season==14)
players14 <- cbPlayers %>% filter(GID %in% games14$GID, Clan=="RUMR")
games13 <- cbGames %>% filter(Season==13)
players13 <- cbPlayers %>% filter(GID %in% games13$GID, Clan=="RUMR")
```

```{r}
playerRates <- players14 %>% group_by(Name) %>% summarise(SurRate = mean(as.numeric(Survived)),WinRate=mean(Result),N=n()) %>% filter(N>5)

playerRates %>%
  ggplot(aes(x=SurRate,y=WinRate,label=Name))+
  scale_color_manual(values=c("TRUE"="cornflowerblue","FALSE"="darkred"))+
  geom_point(aes(colour=SurRate>0.5))+
  geom_text(hjust=1.2,size=2)+
  geom_line(linetype="dotdash",aes(y=0.5, colour="lightgrey",alpha=0.5))+
  geom_line(linetype="dotdash",aes(x=0.5, colour="lightgrey",alpha=0.5))+
  theme_classic()+
  guides(color=guide_legend(title=""))+
  ggtitle("WR vs Survival S14")+
  ylab("Winrate")+
  xlab("Survival Rate")+
  geom_smooth(method='lm', formula= y~x)+
  theme(legend.position = "none")

playerRates <- players13 %>% group_by(Name) %>% summarise(SurRate = mean(as.numeric(Survived)),WinRate=mean(Result),N=n()) %>% filter(N>5)

playerRates %>%
  ggplot(aes(x=SurRate,y=WinRate,label=Name))+
  scale_color_manual(values=c("TRUE"="cornflowerblue","FALSE"="darkred"))+
  geom_point(aes(colour=SurRate>0.5))+
  geom_text(hjust=1.2,size=2)+
  geom_line(linetype="dotdash",aes(y=0.5, colour="lightgrey",alpha=0.5))+
  geom_line(linetype="dotdash",aes(x=0.5, colour="lightgrey",alpha=0.5))+
  theme_classic()+
  guides(color=guide_legend(title=""))+
  ggtitle("WR vs Survival S13")+
  ylab("Winrate")+
  xlab("Survival Rate")+
  geom_smooth(method='lm', formula= y~x)+
  theme(legend.position = "none")
```

```{r}
playerRates <- players14 %>% group_by(Name,Result) %>% summarise(SurRate = mean(as.numeric(Survived)),N=n(),.groups = 'drop') %>% filter(N>5)

playerRates %>%
  ggplot(aes(x=Name,y=SurRate,fill=SurRate>0.5,label=Name))+
  scale_fill_manual(values=c("TRUE"="cornflowerblue","FALSE"="darkred"))+
  geom_col()+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  guides(color=guide_legend(title=""))+
  ggtitle("Survival faceted by Winrate S14")+
  ylab("Survival Rate")+
  xlab("")+
  theme(legend.position = "none")+
  facet_grid(vars(Result))

playerRates <- players13 %>% group_by(Name,Result) %>% summarise(SurRate = mean(as.numeric(Survived)),N=n(),.groups = 'drop') %>% filter(N>5)

playerRates %>%
  ggplot(aes(x=Name,y=SurRate,fill=SurRate>0.5,label=Name))+
  scale_fill_manual(values=c("TRUE"="cornflowerblue","FALSE"="darkred"))+
  geom_col()+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  guides(color=guide_legend(title=""))+
  ggtitle("Survival faceted by Winrate S13")+
  ylab("Survival Rate")+
  xlab("")+
  theme(legend.position = "none")+
  facet_grid(vars(Result))
```

```{r}
playerRates <- players14 %>% group_by(Name,Class) %>% summarise(SurRate = mean(as.numeric(Survived)),WinRate=mean(Result),N=n(),.groups="drop") %>% filter(N>5)

playerRates %>%
  ggplot(aes(x=SurRate,y=WinRate,label=Name))+
  scale_color_manual(values=c("TRUE"="cornflowerblue","FALSE"="darkred"))+
  geom_point()+
  geom_text(hjust=1.2,size=2)+
  geom_line(linetype="dotdash",aes(y=0.5, colour="lightgrey",alpha=0.5))+
  geom_line(linetype="dotdash",aes(x=0.5, colour="lightgrey",alpha=0.5))+
  theme_classic()+
  guides(color=guide_legend(title=""))+
  ggtitle("Winrate and Survival by Class S14")+
  ylab("Winrate")+
  xlab("Survival Rate")+
  theme(legend.position = "none")+
  geom_smooth(method='lm', formula= y~x)+
  facet_wrap(vars(Class))

playerRates <- players13 %>% group_by(Name,Class) %>% summarise(SurRate = mean(as.numeric(Survived)),WinRate=mean(Result),N=n(),.groups="drop") %>% filter(N>5)

playerRates %>%
  ggplot(aes(x=SurRate,y=WinRate,label=Name))+
  scale_color_manual(values=c("TRUE"="cornflowerblue","FALSE"="darkred"))+
  geom_point()+
  geom_text(hjust=1.2,size=2)+
  geom_line(linetype="dotdash",aes(y=0.5, colour="lightgrey",alpha=0.5))+
  geom_line(linetype="dotdash",aes(x=0.5, colour="lightgrey",alpha=0.5))+
  theme_classic()+
  guides(color=guide_legend(title=""))+
  ggtitle("Winrate and Survival by Class S13")+
  ylab("Winrate")+
  xlab("Survival Rate")+
  theme(legend.position = "none")+
  geom_smooth(method='lm', formula= y~x)+
  facet_wrap(vars(Class))
```

What is abundantly clear is that the Tier 6 season (14) was a bit brutal - almost certainly far more results based on kills rather than points compared to T10 seasons. Correlations are not dissimilar to what we see currently, where it is almost certainly more valuable to create value rather than strictly survive. DDs have the flattest fit line almost across the board, which really seems to support this conclusion for that class - taking no risky plays generates low value, but taking too many will average out to low value. 

I can take some solace in the fact that over larger sample sizes my survival rates aren't atrocious, but I should certainly keep them in mind nonetheless.

It doesn't look like there is much to be drawn in the way of conclusions here: survival is clearly somewhat correlated with winning, but this is inevitable since everyone dying is a loss. Without being able to see how long they survived for, or how much value they created before dying, I'm not sure there is anything useful here, outside of seeing some consistent low outlier individuals who might be worth closer inspection.

Like me. I die a lot.

An afterthought: I meant to look at average players left alive by game result.

```{r}
players %>% group_by(GID,Result) %>% summarise(Sur=mean(Survived),.groups="drop") %>%
  ggplot(aes(x=Sur))+
  geom_histogram(stat="bin",bins=10)+
  facet_grid(vars(Result))+
  ylab("Count")+
  xlab("% Team Surviving")+
  ggtitle("Team Survival Percentages S15")+
  geom_density(aes(y=..density..*5),colour="Red")
players14 %>% group_by(GID,Result) %>% summarise(Sur=mean(Survived),.groups="drop") %>%
  ggplot(aes(x=Sur))+
  geom_histogram(stat="bin",bins=10)+
  facet_grid(vars(Result))+
  ggtitle("Team Survival Percentages S14")+
  ylab("Count")+
  xlab("% Team Surviving")+
  geom_density(aes(y=..density..*15),colour="Red")
players13 %>% group_by(GID,Result) %>% summarise(Sur=mean(Survived),.groups="drop") %>%
  ggplot(aes(x=Sur))+
  geom_histogram(stat="bin",bins=10)+
  facet_grid(vars(Result))+
  ggtitle("Team Survival Percentages S13")+
  ylab("Count")+
  xlab("% Team Surviving")+
  geom_density(aes(y=..density..*10),colour="Red")
```

This just repeats my observation before: the tier 6 season was absolutely brutal, and tier 10 games are much more likely to be decided on points. This should probably influence how we intend to play maps: reposition to shore up when we have a cap advantage, and try to work out where we can get that. While kills/losses are huge point swings, if games are going to points or timer, they aren't everything by a longshot.